---
title: "DETAILED REPORT OF BINARY ANALYSES"
author: "Pinheiro R.B.P., Felix G.M.F., and Lewinsohn T.M."
date: "July 2022"
output: pdf_document
toc: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(fig.height = 4, fig.width = 8)
library(here)
knitr::opts_knit$set(root.dir = here())
```

```{r include=FALSE}
library(FSA)
library(stringr)
library(bipartite)
options(digits=3)
```

```{r include=FALSE}
## Results to be analyzed
# If results are not aggregate, define CODES and source line 29 (remove the #).
# If results are already ready, just define CODES for an existing output.
CODES=c("set1","set2","set3")
NAME_OUTPUT=paste(CODES,collapse = "_")#(NAME_OUTPUT in aggregation_results.R)
# Minimum size in each dimension
MIN=0
#source("aggregation_results.R")
# Colors
COLORS1=c(1,"#1E46D9","#F80631","#00D400","#FFA100","#7A05C5")
COLORS2=c("#AF2E45","#ffd143","#353082","#4EA02A")

```

```{r results="hide"}

TABLE_RESULTS=read.table(paste("results/",NAME_OUTPUT,".txt",sep=""), sep="\t", header=T, row.names = 1)
X=read.table(paste("results/","NETDATA_",NAME_OUTPUT,".txt",sep=""), sep="\t", header=T, row.names = 1)
rownames(TABLE_RESULTS)=X$ID
load(paste("results/MODULES_",NAME_OUTPUT,".RData",sep=""))
# File IDs = list of network ids to be analyzed
ids=X$ID
L=length(ids)
# minimum size in each dimension
min_dim=TABLE_RESULTS$rows>=MIN&TABLE_RESULTS$cols>=MIN
#loading networks
NETS=list()
```

# 1. Dataset

Binary networks formed by four kinds of species interaction, comprising almost all the networks analyzed by:  

1. J. Bascompte, P. Jordano, C. J. Melian, J. M. Olesen, The nested assembly of plant-animal mutualistic networks. Proc. Natl. Acad. Sci. 100, 9383–9387 (2003).
2. 	P. Jordano, J. Bascompte, J. M. Olesen, Invariant properties in coevolutionary networks of plant-animal interactions. Ecol. Lett. 6, 69–81 (2003).
3. 	J. M. Olesen, J. Bascompte, Y. L. Dupont, P. Jordano, The modularity of pollination networks. Proc. Natl. Acad. Sci. 104, 19891–19896 (2007).
4. 	M. A. Fortuna, et al., Nestedness versus modularity in ecological networks: two sides of the same coin? J. Anim. Ecol. 79, 811–817 (2010).
5. 	E. Thebault, C. Fontaine, Stability of ecological communities and the architecture of mutualistic and trophic networks. Science (80-. ). 329, 853–856 (2010).


```{r}
print(paste("Networks in the dataset:", L))
knitr::kable(X[order(X$ID),c(1,2,3,5,6)], format = "simple",row.names = F)
barplot(table(TABLE_RESULTS$interaction_type), col=COLORS2[1:4],names.arg = c("Seed dispersion","Plant-Herbivore","Host-Parasite","Pollination"),space = .5,cex.names = .9, main= "Networks by kind of interaction")
text(y=10, x=c(1,2.5,4,5.5),table(TABLE_RESULTS$interaction_type), cex=1.5, col="white")


```

## Warnings (matrix loadings) 

```{r}
for (i in 1:L){
  if(str_detect(X$File[i],pattern = ".txt")){ #txt files
    NETS[[ids[i]]]<-read.table(paste("networks/",X$File[i],sep = ""))}
  if(str_detect(X$File[i],pattern = ".csv")){ #csv files
    NETS[[ids[i]]]<-read.table(paste("networks/",X$File[i],sep = ""), sep = ",")}
  #empty rows and columns
  Rs=rowSums(NETS[[i]])
  Cs=colSums(NETS[[i]])
  if(any(Rs==0)){print(paste(ids[[i]], "had", sum(Rs==0),"empty rows"))}
  if(any(Cs==0)){print(paste(ids[[i]], "had", sum(Cs==0),"empty columns"))}
  NETS[[i]]=NETS[[i]][Rs>0,Cs>0]
  #binary matrix
  if(any(NETS[[ids[i]]]!=0&NETS[[ids[i]]]!=1)){
    print(paste(ids[i], "was not binary"))
    NETS[[ids[i]]][NETS[[ids[i]]]>0]=1
    }
}
```

# 2. Overall nestedness vs. Modularity

Binary nestedness: we applied the NODF index (Almeida-Neto et al. 2008).

Binary modularity: we calculated Barber's modularity index (Barber 2007) using the LPA and the DIRT LPA algorithms (Beckett 2016). 

For too large networks (> 600 species), it was computational impracticable to apply the DIRT LPA algorithm, so we applied the faster, although a bit less effective, LPA.

## List of large binary networks (>600 species)

```{r}
knitr::kable(X[(X$rows+X$cols)>600,c(1,2,3,5,6)], format = "simple", row.names = F)
```

## Plot and correlation test

```{r}
par(mar=c(4,4,0.3,0))
plot(NODF~modularity, pch=16, col=COLORS2[as.factor(interaction_type)], ylim=c(0,100), ylab="nestedness", xlab="modularity", data=TABLE_RESULTS)
legend(x =0.7 ,y=100, legend=c("Seed dispersion","Plant-Herbivore","Host-Parasite","Pollination"),col=COLORS2,pch=16)
```

Testing the correlation between modularity and nestedness:

```{r}
cor.test(TABLE_RESULTS$modularity, TABLE_RESULTS$NODF, method="spearman")
```

# 3. Correlations and comparisons

## 3.1. Connectance:

```{r}
par(mar=c(4,4,0.3,0))
plot(NODF~modularity, pch=16, col=COLORS2[as.factor(interaction_type)], ylim=c(0,100), ylab="nestedness", xlab="modularity", data=TABLE_RESULTS, cex=X$connectance*3)
legend(x =0.7 ,y=100, legend=c("Seed dispersion","Plant-Herbivore","Host-Parasite","Pollination"),col=COLORS2,pch=16)
```

Correlation between modularity and connectance:

```{r}
cor.test(TABLE_RESULTS$modularity, TABLE_RESULTS$connectance, method="spearman")
```

Correlation between nestedness and connectance:

```{r}
cor.test(TABLE_RESULTS$NODF, TABLE_RESULTS$connectance, method="spearman")
```

## 3.2. Network size:

Size was log transformed (size = log(number of species))

```{r}
par(mar=c(4,4,0.3,0))
plot(NODF~modularity, pch=16, col=COLORS2[as.factor(interaction_type)], ylim=c(0,100), ylab="nestedness", xlab="modularity", data=TABLE_RESULTS, cex=log(X$rows+X$cols)/3)
legend(x =0.7 ,y=100, legend=c("Seed dispersion","Plant-Herbivore","Host-Parasite","Pollination"),col=COLORS2,pch=16)
```

Correlation between size and connectance:

```{r}
cor.test((X$rows+X$cols), TABLE_RESULTS$connectance, method="spearman")
```

## 3.3. Nestedness vs. kind of interaction

```{r}
COLORS3=c("#88CC88","#55AA55","#2D882D","#116611","#004400")
```

```{r}
par(mar=c(4,4,0.3,0))
boxplot(NODF~interaction_type, data=TABLE_RESULTS, ylab="nestedness", xlab=NA, names=c("Seed dispersion","Plant-Herbivore","Host-Parasite","Pollination"), pch=16,col=COLORS2[1:4])
```

## 3.4. Modularity vs. kind of interaction

```{r}
par(mar=c(4,4,0.3,0))
boxplot(modularity~interaction_type, data=TABLE_RESULTS, ylab="modularity", xlab=NA, names=c("Seed dispersion","Plant-Herbivore","Host-Parasite","Pollination"), pch=16,col=COLORS2[1:4])
```


# 4. Distortions on null model connectance

We used a null model that forces the randomized matrix to keep the dimensions. It is based on the method developed by Vázquez (2007). However, because of conserving the dimensions, null models are not able to conserve connectance when the number of interactions divided by network size (linkage density) is lower than 1.

Here, to illustrate this problem, we built null models for all networks, regardless of linkage density. Then, we divided the average connectance in the null matrices by the connectance in the observed matrix.

Notice that, in networks with linkage density lower than 1, connectance in the null matrices is often higher than in the observed matrix. This problems never occur for matrices with linkage density equal or higher than 1.

```{r, results="hide"}
linkage_density=(TABLE_RESULTS$nint/(TABLE_RESULTS$rows+TABLE_RESULTS$cols))

```

```{r}
par(mar=c(4,4,2,0))
plot(TABLE_RESULTS$mean_equi_nullconnec_P[min_dim]~linkage_density[min_dim], pch=16, ylab="Connectance: null model / observed ", xlab="Linkage density", main="Equiprobable null model", xlim=c(0,9))
abline(v=1, col="gray50", lty=2)
text(x=1,y=1.25,"Linkage density = 1", col="gray50",pos = 4)

plot(TABLE_RESULTS$mean_prop_nullconnec_P[min_dim]~linkage_density[min_dim], pch=16, ylab="Connectance: null model / observed", xlab="Linkage density", main="Proportional null model",xlim=c(0,9))
abline(v=1, col="gray50", lty=2)
text(x=1,y=1.25,"Linkage density = 1", col="gray50",pos = 4)
```

## 4.1. Networks with linkage density < 1

This is the list of networks in our dataset with likage density lower than 1.

```{r}
X1=X[linkage_density<1&min_dim,c(1,2,3,5,6)]
knitr::kable(X1[order(X1$ID),], format = "simple", row.names = F)
```

Notice that these networks correspond to the 9 networks with higher modularity:

```{r}
par(mar=c(4,4,0.3,0))
plot(NODF~modularity, pch=16, col=COLORS2[as.factor(interaction_type)], ylim=c(0,100), ylab="nestedness", xlab="modularity", data=TABLE_RESULTS)
points(NODF~modularity, pch=1, col=COLORS1[3], ylim=c(0,100), ylab="nestedness", xlab="modularity", data=TABLE_RESULTS[linkage_density<1,], cex=2)
legend(x =0.7 ,y=100, legend=c("Seed dispersion","Plant-Herbivore","Host-Parasite","Pollination","linkage density <1"), col=c(COLORS2,COLORS1[3]),pch=c(16,16,16,16,1),pt.cex = c(1,1,1,1,2))
```


# 5. Modularity and nestedness significance

```{r results="hide"}
criterium1=(linkage_density>=1&min_dim)
```

For these analysis we only included networks with linkage density equal or higher to 1.

Null models in our study are composed of 105 randomized matrices each. This is not a very high number of matrices, but for computational limitations it was impracticable to produce larger null models. We analyzed a large number of real networks (binary: 142, weighted: 68), some of which including a very large number of species (max= 1881 species).

```{r}
print(paste("Number of remaining networks:", sum(criterium1)))
barplot(table(TABLE_RESULTS$interaction_type[criterium1]), col=COLORS2[1:4],names.arg = c("Seed dispersion","Plant-Herbivore","Host-Parasite","Pollination"),space = .5,cex.names = .9,main= "Networks by kind of interaction")
text(y=7, x=c(1,2.5,4,5.5),table(TABLE_RESULTS$interaction_type[criterium1]), cex=1.5, col="white")


par(mar=c(4,4,0.3,0))
plot(NODF~modularity, pch=16, col=COLORS2[as.factor(interaction_type)], ylim=c(0,100), ylab="nestedness", xlab="modularity", data=TABLE_RESULTS[criterium1,])
legend(x =0.5 ,y=100, legend=c("Seed dispersion","Plant-Herbivore","Host-Parasite","Pollination"),col=COLORS2,pch=16)
```

## 5.1. Significantly nested

Nestedness significance was assessed through comparisons with the equiprobable null model.

```{r}
par(mar=c(4,4,0.3,0))
plot(NODF~modularity, pch=c(1,16)[nodf_equi_sig+1], col=COLORS2[as.factor(interaction_type)], ylim=c(0,100), ylab="nestedness", xlab="modularity", yaxt="n", data=TABLE_RESULTS[criterium1,])
legend(x =0.55 ,y=100, legend=c("Seed dispersion","Plant-Herbivore","Host-Parasite","Pollination"," ", "Significant", "Non-significant"),col=c(COLORS2,NA,1,1),pch=c(16,16,16,16,NA,16,1))
```

```{r, include=F}
net_size=X$rows+X$cols
M1=glm(TABLE_RESULTS$nodf_equi_sig[criterium1]~net_size[criterium1], family="binomial")
C=coef(M1)
```

```{r}
print(paste("Number of significantly nested networks:", sum(TABLE_RESULTS$nodf_equi_sig[criterium1])))
par(mar=c(4,4,0.3,0))
net_size=TABLE_RESULTS$cols+TABLE_RESULTS$rows
plot(nodf_equi_sig~net_size[criterium1],data=TABLE_RESULTS[criterium1,], pch=16, ylab="Nestedness significance", xlab="Network size")
curve((exp(C[[1]]+C[[2]]*x)/(1+exp(C[[1]]+C[[2]]*x))), add=T)
```

All the networks with more than 34 species were significantly nested.

## 5.2. Significantly modular

Modularity significance was assessed through comparisons with the proportional null model.

```{r}
par(mar=c(4,4,0.3,0))
plot(NODF~modularity, pch=c(1,16)[mod_prop_sig+1], col=COLORS2[as.factor(interaction_type)], ylim=c(0,100), ylab="nestedness", xlab="modularity", data=TABLE_RESULTS[criterium1,])
legend(x =0.55 ,y=100, legend=c("Seed dispersion","Plant-Herbivore","Host-Parasite","Pollination"," ", "Significant", "Non-significant"),col=c(COLORS2,NA,1,1),pch=c(16,16,16,16,NA,16,1))
```

```{r, include=F}
M1=glm(TABLE_RESULTS$mod_prop_sig[criterium1]~net_size[criterium1], family="binomial")
C=coef(M1)
```

```{r}
print(paste("Number of significantly modular networks:", sum(TABLE_RESULTS$mod_prop_sig[criterium1])))
par(mar=c(4,4,0.3,0))
net_size=TABLE_RESULTS$cols+TABLE_RESULTS$rows
plot(mod_prop_sig~net_size[criterium1],data=TABLE_RESULTS[criterium1,], pch=16, ylab="Modularity significance", xlab="Network size", yaxt="n")
curve((exp(C[[1]]+C[[2]]*x)/(1+exp(C[[1]]+C[[2]]*x))), add=T)
```

Almost all the networks (except from two) with more than 160 species were significantly modular. 

# 6. Nestedness SM and DM

For significant modular networks we calculated nestedness between species belonging to the same module (N~SM~) and nestedness between species in different modules (N~DM~) (Flores et al. 2013, Pinheiro et al. 2019, Felix et al. 2022).

```{r, results="hide"}
criterium2=TABLE_RESULTS$mod_prop_sig&criterium1
```

```{r}
print(paste("Modular networks:", sum(criterium2)))
barplot(table(TABLE_RESULTS$interaction_type[criterium2]), col=COLORS2[1:4],names.arg = c("Seed dispersion","Plant-Herbivore","Host-Parasite","Pollination"),space = .5,cex.names = .9,main= "Modular networks by kind of interaction")
text(y=3, x=c(1,2.5,4,5.5),table(TABLE_RESULTS$interaction_type[criterium2]), cex=1.5, col="white")
```

```{r}
par(mar=c(4,4,0.3,0))
plot(NODF~modularity, pch=NA, ylim=c(0,100), ylab="nestedness", xlab="modularity", data=TABLE_RESULTS[criterium2,])
points(NODF_SM~modularity, pch=16, col=COLORS1[2], data=TABLE_RESULTS[criterium2,])
points(NODF_DM~modularity, pch=16, col=COLORS1[3], data=TABLE_RESULTS[criterium2,])
points(NODF~modularity, pch=16, col=COLORS1[1], data=TABLE_RESULTS[criterium2,])
legend(x =0.55 ,y=100, legend=c("Overall","SM","DM"),col=COLORS1[1:3],pch=16)

```

```{r}
par(mar=c(4,4,0.3,0))
plot(NODF~modularity, pch=NA, ylim=c(0,100), ylab="nestedness", xlab="modularity", data=TABLE_RESULTS[criterium2,])
arrows(x0=TABLE_RESULTS$modularity[criterium2],y0=TABLE_RESULTS$NODF[criterium2],y1=TABLE_RESULTS$NODF_SM[criterium2],col=COLORS1[2],length = 0)
arrows(x0=TABLE_RESULTS$modularity[criterium2],y0=TABLE_RESULTS$NODF[criterium2],y1=TABLE_RESULTS$NODF_DM[criterium2],col=COLORS1[3],length = 0)
points(NODF~modularity, pch=16, col=COLORS1[1], data=TABLE_RESULTS[criterium2,])
points(NODF_SM~modularity, pch=16, col=COLORS1[2], data=TABLE_RESULTS[criterium2,])
points(NODF_DM~modularity, pch=16, col=COLORS1[3], data=TABLE_RESULTS[criterium2,])
legend(x =0.55 ,y=100, legend=c("Overall","SM","DM"),col=COLORS1[1:3],pch=16)
```

I applied local regressions (R function loess), very smoothly, to see the general trend of NODF~O~, NODF~SM~ and NODF~DM~, with increasing modularity.


```{r}
par(mar=c(4,4,0.3,0))
plot(NODF~modularity, pch=NA, ylim=c(0,100), ylab="nestedness", xlab="modularity", data=TABLE_RESULTS[criterium2,])
arrows(x0=TABLE_RESULTS$modularity[criterium2],y0=TABLE_RESULTS$NODF[criterium2],y1=TABLE_RESULTS$NODF_SM[criterium2],col=COLORS1[2],length = 0)
arrows(x0=TABLE_RESULTS$modularity[criterium2],y0=TABLE_RESULTS$NODF[criterium2],y1=TABLE_RESULTS$NODF_DM[criterium2],col=COLORS1[3],length = 0)
points(NODF~modularity, pch=16, col=COLORS1[1], data=TABLE_RESULTS[criterium2,])
points(NODF_SM~modularity, pch=16, col=COLORS1[2], data=TABLE_RESULTS[criterium2,])
points(NODF_DM~modularity, pch=16, col=COLORS1[3], data=TABLE_RESULTS[criterium2,])
pred1=predict(loess(TABLE_RESULTS[criterium2,]$NODF~TABLE_RESULTS[criterium2,]$modularity,span = 1), type="response", newdata = seq(min(TABLE_RESULTS[criterium2,]$modularity),max(TABLE_RESULTS[criterium2,]$modularity),length.out=10000))
lines(pred1~seq(min(TABLE_RESULTS[criterium2,]$modularity),max(TABLE_RESULTS[criterium2,]$modularity),length.out=10000), lty=2)
pred2=predict(loess(TABLE_RESULTS[criterium2,]$NODF_SM~TABLE_RESULTS[criterium2,]$modularity,span = 1), type="response", newdata = seq(min(TABLE_RESULTS[criterium2,]$modularity),max(TABLE_RESULTS[criterium2,]$modularity),length.out=10000))
lines(pred2~seq(min(TABLE_RESULTS[criterium2,]$modularity),max(TABLE_RESULTS[criterium2,]$modularity),length.out=10000), lty=2, col=COLORS1[2])
pred3=predict(loess(TABLE_RESULTS[criterium2,]$NODF_DM~TABLE_RESULTS[criterium2,]$modularity,span = 1), type="response", newdata = seq(min(TABLE_RESULTS[criterium2,]$modularity),max(TABLE_RESULTS[criterium2,]$modularity),length.out=10000))
lines(pred3~seq(min(TABLE_RESULTS[criterium2,]$modularity),max(TABLE_RESULTS[criterium2,]$modularity),length.out=10000), lty=2, col=COLORS1[3])
legend(x =0.55 ,y=100, legend=c("Overall","SM","DM"),col=COLORS1[1:3],pch=16)
rm(pred1, pred2, pred3)
```

# 7. Significance of nestedness SM

We used restricted null models (models that conserve the modular structure on the networks) to test N~SM~ significance.

Brackets represent 95% of values in the null model.

```{r}
par(mar=c(4,4,0.3,0))
plot(equi_null_SM_mean~modularity, pch=NA, col=COLORS1[1], ylim=c(0,90), ylab="nestedness", xlab="modularity", data=TABLE_RESULTS[criterium2,])
points(NODF_SM~modularity, pch=16, col=COLORS1[2], data=TABLE_RESULTS[criterium2,])
points(equi_null_SM_mean~modularity, pch=16, cex=.7, col="grey30", data=TABLE_RESULTS[criterium2,])
arrows(x0=TABLE_RESULTS$modularity[criterium2],y0=TABLE_RESULTS$equi_null_SM_mean[criterium2],y1=TABLE_RESULTS$equi_SM_ci_up[criterium2],col="grey30",length = .05,angle=90)
arrows(x0=TABLE_RESULTS$modularity[criterium2],y0=TABLE_RESULTS$equi_null_SM_mean[criterium2],y1=TABLE_RESULTS$equi_SM_ci_down[criterium2],col="grey30",length = .05,angle=90)
legend(x =0.55 ,y=90, legend=c("SM", "Null model"),col=c(COLORS1[2],"grey30"),pch=16)
```

Only a few modular networks do not present significant nestedness SM (red arrows).

```{r}
par(mar=c(4,4,0.3,0))
plot(equi_null_SM_mean~modularity, pch=NA, col=COLORS1[1], ylim=c(0,90), ylab="nestedness", xlab="modularity", data=TABLE_RESULTS[criterium2,])
points(NODF_SM~modularity, pch=16, col=COLORS1[2], data=TABLE_RESULTS[criterium2,])
points(equi_null_SM_mean~modularity, pch=16, cex=.7, col="grey30", data=TABLE_RESULTS[criterium2,])
arrows(x0=TABLE_RESULTS$modularity[criterium2],y0=TABLE_RESULTS$equi_null_SM_mean[criterium2],y1=TABLE_RESULTS$equi_SM_ci_up[criterium2],col="grey30",length = .05,angle=90)
arrows(x0=TABLE_RESULTS$modularity[criterium2],y0=TABLE_RESULTS$equi_null_SM_mean[criterium2],y1=TABLE_RESULTS$equi_SM_ci_down[criterium2],col="grey30",length = .05,angle=90)
arrows(x0=TABLE_RESULTS$modularity[criterium2&TABLE_RESULTS$topology=="pure modular"],y0=TABLE_RESULTS$equi_null_SM_mean[criterium2&TABLE_RESULTS$topology=="pure modular"]-13,y1=TABLE_RESULTS$equi_null_SM_mean[criterium2&TABLE_RESULTS$topology=="pure modular"]-7, length = .1, lwd=4, col=COLORS1[3])
legend(x =0.55 ,y=90, legend=c("SM", "Null model"),col=c(COLORS1[2],"grey30"),pch=16)
```

# 8. Network topologies

Significance tests:

Equiprobable null model for N~O~ significance.

Proportional null model for modularity significance.

Equiprobable restricted null model for N~SM~ significance.

Topologies:

Nested network: non-significant modularity and significant N~O~

Pure modular network: significant modularity and non-significant N~SM~

Compound topology: singnificant modularity and significant N~SM~

Unstructured: non significant N~O~ and non significant modularity

```{r}
TAB1=table(TABLE_RESULTS$topology[criterium1])
knitr::kable(TAB1,col.names = c("topology","Number of networks"), format = "simple")
```

## 8.1. Topology vs. kind of interaction

```{r}
TAB3=TABLE_RESULTS[criterium1,c(1,8)]
TAB3.2=TABLE_RESULTS[linkage_density<1&min_dim,c(1,8)]
TAB3.2$topology="linkage density <1"
TAB3=rbind(TAB3,TAB3.2)
TAB3=table(TAB3)
TAB3=TAB3[c(3,5,1,4,2),]
knitr::kable(TAB3,col.names = c("seed dispersal","plant-herbivore","host-parasite","pollination"), format = "simple")
```

# 9. Network robustness

Here, we simulate the exclusion of species on one side of the network (row or columns in the matrix) and analyze the secondary extinctions: species on the other size that lose all interactions. We only used the binary matrices for this analysis.

As the measure of robustness we applied the index proposed by Burgos et al. (2007), which is the area under the secondary extinction curve. We restricted this analysis to nested and compound networks, as they represent the large majority of binary networks (124 out of 142). 

## 9.1 Removal of resource species

```{r, include=F}
load("results/robustness_set1_set2_set3.RData")
criterium3=ROB_RESULTS$topology=="nested"|ROB_RESULTS$topology=="compound"
criterium3[!criterium1]=FALSE
```

We removed the resource species (rows): plants in seed dispersal, pollination and plant-herbivore networks, and hosts in host-parasite networks.

For all nested or compound networks, we measured robustness to resource species removals (1) based on species degrees (more connected to less connected) and (2) randomly. For each network, we made 1000 random extinction sequences and calculated the average robustness.

```{r, fig.height = 5, fig.width = 5}
boxplot(ROB_RESULTS$rob_low_deg[criterium3]~ROB_RESULTS$topology[criterium3], col=COLORS2[c(3,1)], xlab="topology", ylab="robustness", main="removals by degrees", ylim=c(0,1))
```

We performed a non-parametrical Mann-Whitney test to compare the robustness of nested and compound binary networks:

```{r}
wilcox.test(ROB_RESULTS$rob_low_deg[criterium3&ROB_RESULTS$topology=="nested"],ROB_RESULTS$rob_low_deg[criterium3&ROB_RESULTS$topology=="compound"])

```

Nested networks were significantly more robust to removals of resource species following decreasing degrees than networks with compound topologies.

```{r, fig.height = 5, fig.width = 5}
boxplot(ROB_RESULTS$rob_low_ran[criterium3]~ROB_RESULTS$topology[criterium3], col=COLORS2[c(3,1)], xlab="topology", ylab="average robustness (1000 extinction sequences)", main="random removals",ylim=c(0,1))
```

We performed a non-parametrical Mann-Whitney test to compare the robustness of nested and compound binary networks:

```{r}
wilcox.test(ROB_RESULTS$rob_low_ran[criterium3&ROB_RESULTS$topology=="nested"],ROB_RESULTS$rob_low_ran[criterium3&ROB_RESULTS$topology=="compound"])

```
We did not find statistically significant differences in the robustness of nested and compound networks to random sequences of resource species removals.
 
 \
 
For networks with compound topologies (n=45) we compared the robustness following two different rules for the sequences of species removals. For each network, using each rule, we made 1000 different sequences of extinctions.

In the first rule, we maximally delay the exclusion of modules (removal of all species of the module). For this, in each round, the resource species to be removed is randomly selected from the module or modules with more remaining resources.

Imagine a network with 15 resource species, distributed among 3 modules: A, B and C. Initially, A has 8 resources, B has 4 and C has 3. This is a possible sequence of removals following rule 1.

remaining species in each module -> action

A=8 B=4 C=3 -> remove a random species from A \
A=7 B=4 C=3 -> remove a random species from A \
A=6 B=4 C=3 -> remove a random species from A \
A=5 B=4 C=3 -> remove a random species from A \
A=4 B=4 C=3 -> remove a random species from A or B \
A=4 B=3 C=3 -> remove a random species from A \
A=3 B=3 C=3 -> remove a random species from A, B or C \
A=2 B=3 C=3 -> remove a random species from B or C \
A=2 B=3 C=2 -> remove a random species from B \
A=2 B=2 C=2 -> remove a random species from A, B or C \
A=2 B=1 C=2 -> remove a random species from A or C \
A=2 B=1 C=1 -> remove a random species from A \
A=1 B=1 C=1 -> remove a random species from A, B or C \
A=1 B=0 C=1 -> remove a random species from A or C \
A=0 B=0 C=1 -> remove a random species from C \
A=0 B=0 C=0 - End of extinction sequence \

The second rule for species removals aims at hastening the exclusion of modules.Therefore, for each sequence of removals, we randomly define the order of modules to be tackled. All modules have the same chance of being selected in any position. Then, we proceed by removing species from the first defined module, until it is fully excluded, then we move to the next and so on.

Imagine a network with 15 resource species, distributed among 3 modules: A, B and C. Initially, A has 8 resources, B has 4 and C has 3. First we define a random order of modules: B - A - C. In the first 4 rounds, random species of module B would be removed, then in the next 8 rounds, the species of A are removed, and finally, in the last 3 rounds, the species of C are removed. 


```{r, include=F}
compound=criterium3&ROB_RESULTS$topology=="compound"
statistic=numeric()
pvalue=numeric()
for (JJJ in 1:142){
  if(!is.null(COMPOUND1ROB_RESULTS_L[[JJJ]])){
    TESTE=wilcox.test(COMPOUND1ROB_RESULTS_L[[JJJ]],COMPOUND2ROB_RESULTS_L[[JJJ]])
    statistic[JJJ]=TESTE$statistic
    pvalue[JJJ]=TESTE$p.value
  }
}
```

```{r, fig.height = 5}
plot(NA, ylim=c(0.5,1), xlim=c(.5,2.5), ylab="average robustness (1000 extinction sequences)",xaxt='n', xlab=NA)
Axis(side=1,at = c(1,2), labels = c("delaying module exclusions","hastening module exclusions"))
for(JJJ in 1:nrow(ROB_RESULTS)){
  if (!is.null(COMPOUND1ROB_RESULTS_L[[JJJ]])){
    lines(x=c(1,2),y=ROB_RESULTS[JJJ,c(10,9)], lwd=1.5, col="gray50")
    points(x=c(1,2),y=ROB_RESULTS[JJJ,c(10,9)], pch=16, col=COLORS2[3])
  }
}
boxplot(ROB_RESULTS[,c(10,9)], add = T,at = c(.7,2.3),boxwex=.3,varwidth = FALSE, col = COLORS2[3], lwd=2, axes=F)
```

We performed non-parametrical Mann-Whitney tests for each compound network to compare its robustness to sequences of removals following the two rules:


```{r}
knitr::kable(data.frame(networks=ids[compound],statistic=statistic[compound], pvalue=format(pvalue[compound],scientific = T)), format = "simple",row.names = F)
```

All networks were significantly more robust to sequences of removals that delay the exclusion of modules than to sequences of removals that hasten the exclusion of modules.


## 9.1 Removal of consumer species

We removed the consumer species (rows): animals in seed dispersal, pollination and plant-herbivore networks, and parasites in host-parasite networks.

For all nested or compound networks, we measured robustness to consumer species removals (1) based on species degrees (more connected to less connected) and (2) randomly. For each network, we made 1000 random extinction sequences and calculated the average robustness.

```{r, fig.height = 5, fig.width = 5}
boxplot(ROB_RESULTS$rob_high_deg[criterium3]~ROB_RESULTS$topology[criterium3], col=COLORS2[c(3,1)], xlab="topology", ylab="robustness", main="removals by degrees", ylim=c(0,1))
```

We performed a non-parametrical Mann-Whitney test to compare the robustness of nested and compound binary networks:
  
```{r}
wilcox.test(ROB_RESULTS$rob_high_deg[criterium3&ROB_RESULTS$topology=="nested"],ROB_RESULTS$rob_high_deg[criterium3&ROB_RESULTS$topology=="compound"])

```
We did not find statistically significant differences in the robustness of nested and compound networks to sequences of consumer species removals by decreasing degrees.


```{r, fig.height = 5, fig.width = 5}
boxplot(ROB_RESULTS$rob_high_ran[criterium3]~ROB_RESULTS$topology[criterium3], col=COLORS2[c(3,1)], xlab="topology", ylab="mean robustness (1000 extinction sequences)", main="random removals", ylim=c(0,1))
```

We performed a non-parametrical Mann-Whitney test to compare the robustness of nested and compound binary networks:

```{r}
wilcox.test(ROB_RESULTS$rob_high_ran[criterium3&ROB_RESULTS$topology=="nested"],ROB_RESULTS$rob_high_ran[criterium3&ROB_RESULTS$topology=="compound"])

```

Networks with compound topologies were significantly more robust to random removals of consumer species than nested networks.

\

For networks with compound topologies (n=45) we compared the robustness following two different rules for the sequences of species removals. For each network, using each rule, we made 1000 different sequences of extinctions.

In the first rule, we maximally delay the exclusion of modules (removal of all species of the module). For this, in each round, the consumer species to be removed is randomly selected from the module or modules with more remaining resources.

The second rule for species removals aims at hastening the exclusion of modules.Therefore, for each sequence of removals, we randomly define the order of modules to be tackled. All modules have the same chance of being selected in any position. Then, we proceed by removing species from the first defined module, until it is fully excluded, then we move to the next and so on.


```{r, include=F}
compound=criterium3&ROB_RESULTS$topology=="compound"
statistic=numeric()
pvalue=numeric()
for (JJJ in 1:142){
  if(!is.null(COMPOUND1ROB_RESULTS_H[[JJJ]])){
    TESTE=wilcox.test(COMPOUND1ROB_RESULTS_H[[JJJ]],COMPOUND2ROB_RESULTS_H[[JJJ]])
    statistic[JJJ]=TESTE$statistic
    pvalue[JJJ]=TESTE$p.value
  }
}
```

```{r, fig.height = 5}
plot(NA, ylim=c(0.5,1), xlim=c(.5,2.5), ylab="mean robustness (1000 extinction sequences)",xaxt='n', xlab=NA)
Axis(side=1,at = c(1,2), labels = c("delaying module exclusions","hastening module exclusions"))
for(JJJ in 1:nrow(ROB_RESULTS)){
  if (!is.null(COMPOUND1ROB_RESULTS_H[[JJJ]])){
    lines(x=c(1,2),y=ROB_RESULTS[JJJ,c(14,13)], lwd=1.5, col="gray50")
    points(x=c(1,2),y=ROB_RESULTS[JJJ,c(14,13)], pch=16, col=COLORS2[3])
  }
}
boxplot(ROB_RESULTS[,c(14,13)], add = T,at = c(.7,2.3),boxwex=.3,varwidth = FALSE, col = COLORS2[3], lwd=2, axes=F)
```
We performed non-parametrical Mann-Whitney tests for each compound network to compare its robustness to sequences of removals following the two rules:

```{r}
knitr::kable(data.frame(networks=ids[compound],statistic=statistic[compound], pvalue=format(pvalue[compound],scientific = T)), format = "simple",row.names = F)
```

All networks were significantly more robust to sequences of removals that delay the exclusion of modules than to sequences of removals that hasten the exclusion of modules.

# 10. References

D. P. Vázquez, et al., Species abundance and asymmetric interaction strength in ecological networks. Oikos 116, 1120–1127 (2007).

E. Burgos, et al. Why nestedness in mutualistic networks? Journal of Theoretical Biology 249, 307–313 (2007). 

M. Almeida-Neto, P. Guimarães, P. R. Guimarães, R. D. Loyola, W. Ulrich, A consistent metric for nestedness analysis in ecological systems: reconciling concept and measurement. Oikos 117, 1227–1239 (2008).

C. O. Flores, S. Valverde, J. S. Weitz, Multi-scale structure and geographic drivers of cross-infection within marine bacteria and phages. ISME J. 7, 520–532 (2013).

S. J. Beckett, Improved community detection in weighted bipartite networks. R. Soc. Open Sci. 3, 140536 (2016).

R. B. P. Pinheiro, G. M. F. Felix, C. F. Dormann, M. A. R. Mello, A new model explaining the origin of different topologies in interaction networks. Ecology 100, 1–30 (2019).

G. M. Felix, R. B. P. Pinheiro, R. Poulin, B. R. Krasnov, M. A. R. Mello, The compound topology of host–parasite networks is explained by the integrative hypothesis of specialization. Oikos 2022 (2022).